---
title: "MultiP_Function_ClassifyR"
author: "Andy Tran, Andy Wang & Jamie Mickaill"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    self_contained: true
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
always_allow_html: true
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(MultiAssayExperiment)
  library(tidyverse)
  library(tictoc)
  library(ClassifyR)
  library(curatedTCGAData)
})
```

# Melanoma (SKCM)

## in house data processing

```{r}
dataDirMelanoma = "data/melanomaAssaysNorm.RData"
load(dataDirMelanoma)

#melanomaAssaysNorm <- mergeReplicates(intersectColumns(melanomaAssaysNorm))

melanomaAssaysNorm = melanomaAssaysNorm[,,c("NanoString", "microRNA")]

melanomaAssaysNorm$pathology_N_stage <- as.character(melanomaAssaysNorm$Prim_NStage)
melanomaAssaysNorm$pathology_N_stage[grep("n0", melanomaAssaysNorm$pathology_N_stage, ignore.case = TRUE)] <- "N0"
melanomaAssaysNorm$pathology_N_stage[grep("n1", melanomaAssaysNorm$pathology_N_stage, ignore.case = TRUE)] <- "N1"
melanomaAssaysNorm$pathology_N_stage[grep("n2", melanomaAssaysNorm$pathology_N_stage, ignore.case = TRUE)] <- "N2"
melanomaAssaysNorm$pathology_N_stage[grep("nx", melanomaAssaysNorm$pathology_N_stage, ignore.case = TRUE)] <- NA

melanomaAssaysNorm$pathology_T_stage <- as.character(melanomaAssaysNorm$Prim_Tstage)
melanomaAssaysNorm$pathology_T_stage[grep("t0", melanomaAssaysNorm$pathology_T_stage, ignore.case = TRUE)] <- "T0"
melanomaAssaysNorm$pathology_T_stage[grep("t1", melanomaAssaysNorm$pathology_T_stage, ignore.case = TRUE)] <- "T1"
melanomaAssaysNorm$pathology_T_stage[grep("t2", melanomaAssaysNorm$pathology_T_stage, ignore.case = TRUE)] <- "T2"
melanomaAssaysNorm$pathology_T_stage[grep("t3", melanomaAssaysNorm$pathology_T_stage, ignore.case = TRUE)] <- "T3"
melanomaAssaysNorm$pathology_T_stage[grep("t4", melanomaAssaysNorm$pathology_T_stage, ignore.case = TRUE)] <- "T4"
melanomaAssaysNorm$pathology_T_stage[grep("tx", melanomaAssaysNorm$pathology_T_stage, ignore.case = TRUE)] <- NA

melanomaAssaysNorm$sex <- tolower(as.character(melanomaAssaysNorm$Person_Sex))
melanomaAssaysNorm$ulceration <- tolower(as.character(melanomaAssaysNorm$Prim_Ulc))
melanomaAssaysNorm$Breslow_thickness <- melanomaAssaysNorm$Prim_Breslow
melanomaAssaysNorm$age <- melanomaAssaysNorm$Age_Analysis/365

#melanoma_clinical_features <- c("pathology_N_stage", "pathology_T_stage", "sex", "Breslow_thickness",)
# melanoma_clinical_features <- c("pathology_T_stage", "sex", "Breslow_thickness", "age")
#melanoma_clinical_features <- c("pathology_N_stage", "pathology_T_stage", "sex", "age")
melanoma_clinical_features <- c("pathology_T_stage", "sex", "age")

classes = "class1vs4years" # outcome feature
keep_features <- colnames(colData(melanomaAssaysNorm)) %in% c(melanoma_clinical_features, classes)
colData(melanomaAssaysNorm) <- colData(melanomaAssaysNorm)[,keep_features]
colData(melanomaAssaysNorm) <- na.omit(colData(melanomaAssaysNorm))

for(feature in melanoma_clinical_features){
    if(class(colData(melanomaAssaysNorm)[,feature]) == "character"){
    colData(melanomaAssaysNorm)[,feature] = as.numeric(factor(colData(melanomaAssaysNorm)[,feature]))
    }
}

#Subset complete cases only for all assays
melanomaAssaysNorm = melanomaAssaysNorm[,complete.cases(melanomaAssaysNorm),]

dim(colData(melanomaAssaysNorm))
melanomaAssaysNorm@ExperimentList
```



```{r, eval = TRUE}
SKCM = curatedTCGAData(diseaseCode = "SKCM", 
                       assays = c("RNASeq2GeneNorm", 
                                  "miRNASeqGene"), 
                       version = "2.0.1", 
                       dry.run = FALSE)

SKCM <- mergeReplicates(intersectColumns(SKCM))
#names(SKCM) <- c("RNAseq", "RPPA", "GISTIC")

SKCM = c(SKCM, 
         RNASeq = SKCM@ExperimentList[["SKCM_RNASeq2GeneNorm-20160128"]], 
         mapFrom = "SKCM_RNASeq2GeneNorm-20160128")

SKCM = c(SKCM, 
         miRNA = SKCM@ExperimentList[["SKCM_miRNASeqGene-20160128"]], 
         mapFrom = "SKCM_miRNASeqGene-20160128")

SKCM = SKCM[,,c("RNASeq", "miRNA")]

experiments(SKCM)
dim(colData(SKCM))

## filter to stage 3
met_ind <- which(str_detect(colData(SKCM)$pathologic_stage, "stage iii"))
SKCM <- SKCM[,met_ind,]
dim(colData(SKCM))

good_ind_alive <- which(colData(SKCM)$vital_status == 0 & colData(SKCM)$days_to_last_followup > 4*365 )
good_ind_dead <- which(colData(SKCM)$vital_status == 1 & colData(SKCM)$days_to_death > 4*365 )
good_ind <- c(good_ind_alive, good_ind_dead)
poor_ind <- which(colData(SKCM)$vital_status == 1 & colData(SKCM)$days_to_death < 1*365 )

colData(SKCM)$Outcome <- NA
colData(SKCM)$Outcome[good_ind] <- "Good"
colData(SKCM)$Outcome[poor_ind] <- "Poor"
table(SKCM$Outcome)


#SKCM 
SKCM$pathology_N_stage <- as.character(SKCM$pathology_N_stage)
SKCM$pathology_N_stage[grep("n0", SKCM$pathology_N_stage, ignore.case = TRUE)] <- "N0"
SKCM$pathology_N_stage[grep("n1", SKCM$pathology_N_stage, ignore.case = TRUE)] <- "N1"
SKCM$pathology_N_stage[grep("n2", SKCM$pathology_N_stage, ignore.case = TRUE)] <- "N2"
SKCM$pathology_N_stage[grep("n3", SKCM$pathology_N_stage, ignore.case = TRUE)] <- "N3"
SKCM$pathology_N_stage[grep("nx", SKCM$pathology_N_stage, ignore.case = TRUE)] <- NA

SKCM$pathology_T_stage <- as.character(SKCM$pathology_T_stage)
SKCM$pathology_T_stage[grep("t0", SKCM$pathology_T_stage, ignore.case = TRUE)] <- "T0"
SKCM$pathology_T_stage[grep("t1", SKCM$pathology_T_stage, ignore.case = TRUE)] <- "T1"
SKCM$pathology_T_stage[grep("t2", SKCM$pathology_T_stage, ignore.case = TRUE)] <- "T2"
SKCM$pathology_T_stage[grep("t3", SKCM$pathology_T_stage, ignore.case = TRUE)] <- "T3"
SKCM$pathology_T_stage[grep("t4", SKCM$pathology_T_stage, ignore.case = TRUE)] <- "T4"
SKCM$pathology_T_stage[grep("tx", SKCM$pathology_T_stage, ignore.case = TRUE)] <- NA
SKCM$pathology_T_stage[grep("tis", SKCM$pathology_T_stage, ignore.case = TRUE)] <- NA

SKCM$sex <- SKCM$gender
SKCM$age <- SKCM$years_to_birth
SKCM$ulceration <- SKCM$melanoma_ulceration

#SKCM_clinical_features <- c("pathology_T_stage", "pathology_N_stage", "sex", "Breslow_thickness", "age")
# SKCM_clinical_features <- c("pathology_T_stage", "sex", "Breslow_thickness", "age")
#SKCM_clinical_features <- c("pathology_N_stage", "pathology_T_stage", "sex", "age")
SKCM_clinical_features <- c("pathology_T_stage", "sex", "age")

classes = "Outcome" # outcome feature
keep_features <- colnames(colData(SKCM)) %in% c(SKCM_clinical_features, classes)
colData(SKCM) <- colData(SKCM)[,keep_features]
colData(SKCM) <- na.omit(colData(SKCM))

for(feature in SKCM_clinical_features){
    if(class(colData(SKCM)[,feature]) == "character"){
    colData(SKCM)[,feature] = as.numeric(factor(colData(SKCM)[,feature]))
    }
}

## library size normalisation as well???
normalise <- function(counts){
  libsizes <- colSums(counts)
  size.factors <- libsizes/mean(libsizes)
  normalised <- log2(t(t(counts)/size.factors) + 1)
  return(normalised)
}


RNAseq <- assay(SKCM@ExperimentList[["RNASeq"]])
assay(SKCM@ExperimentList[["RNASeq"]]) <- normalise(RNAseq)
# keep_RNAseq <- which(apply(RNAseq, 1 , sd) > 1 & rowMeans(RNAseq) > 1)
# SKCM <- SKCM[list(RNASeq = keep_RNAseq),,]

miRNA <- assay(SKCM@ExperimentList[["miRNA"]])
assay(SKCM@ExperimentList[["miRNA"]]) <- normalise(miRNA)
# keep_miRNA <- which(apply(miRNA, 1 , sd) > 3 & rowMeans(miRNA) > 3) # medium filter
keep_miRNA <- which(apply(miRNA, 1 , sd) > 5 & rowMeans(miRNA) > 5) # hard filter
SKCM <- SKCM[list(miRNA = keep_miRNA),,]



dim(colData(SKCM))
experiments(SKCM)
```

## filter down to common genes

```{r}
SKCM_rnaseq_genes <- SKCM@ExperimentList[["RNASeq"]]@NAMES
SKCM@ExperimentList[["RNASeq"]] <- SKCM@ExperimentList[["RNASeq"]][order(SKCM_rnaseq_genes),]
SKCM_rnaseq_genes <- SKCM@ExperimentList[["RNASeq"]]@NAMES

inHouse_targeted_genes <- rownames(melanomaAssaysNorm@ExperimentList[["NanoString"]])
melanomaAssaysNorm@ExperimentList[["NanoString"]] <- melanomaAssaysNorm@ExperimentList[["NanoString"]][order(inHouse_targeted_genes),]
inHouse_targeted_genes <- rownames(melanomaAssaysNorm@ExperimentList[["NanoString"]])

common_genes <- intersect(SKCM_rnaseq_genes, inHouse_targeted_genes)

SKCM <- SKCM[list(RNASeq = SKCM_rnaseq_genes %in% common_genes),,]
melanomaAssaysNorm <- melanomaAssaysNorm[list(NanoString = inHouse_targeted_genes %in% common_genes),,]

rownames(melanomaAssaysNorm@ExperimentList[["NanoString"]]) <- sub("-", ".", rownames(melanomaAssaysNorm@ExperimentList[["NanoString"]]), fixed = TRUE)
rownames(SKCM@ExperimentList[["RNASeq"]]) <- sub("-", ".", rownames(SKCM@ExperimentList[["RNASeq"]]), fixed = TRUE)

SKCM@ExperimentList
melanomaAssaysNorm@ExperimentList
```

## filter down to common miRNAs

```{r}
SKCM_miRs <- SKCM@ExperimentList[["miRNA"]]@NAMES
inHouse_miRs <- rownames(melanomaAssaysNorm@ExperimentList[["microRNA"]])
inHouse_miRs <- sub("R", "r", inHouse_miRs)
rownames(melanomaAssaysNorm@ExperimentList[["microRNA"]]) <- inHouse_miRs

common_miRs <- intersect(SKCM_miRs, inHouse_miRs)

SKCM <- SKCM[list(miRNA = SKCM_miRs %in% common_miRs),,]

melanomaAssaysNorm <- melanomaAssaysNorm[list(microRNA = inHouse_miRs %in% common_miRs),,]

rownames(melanomaAssaysNorm@ExperimentList[["microRNA"]]) <- sub("-", ".", rownames(melanomaAssaysNorm@ExperimentList[["microRNA"]]), fixed = TRUE)

rownames(SKCM@ExperimentList[["miRNA"]]) <- sub("-", ".", rownames(SKCM@ExperimentList[["miRNA"]]), fixed = TRUE)

SKCM@ExperimentList
melanomaAssaysNorm@ExperimentList
```



## put data on same scale (RNAseq only)

```{r}
par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["RNASeq"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["NanoString"]]))
```


```{r}
SKCM_RNA_t <- t(assay(SKCM@ExperimentList[["RNASeq"]]))
MIA_RNA_t <- t(assay(melanomaAssaysNorm@ExperimentList[["NanoString"]]))
identical(colnames(MIA_RNA_t), colnames(SKCM_RNA_t))


SKCM_RNA_t <- log2(SKCM_RNA_t + 1)
MIA_RNA_t <- log2(MIA_RNA_t - min(MIA_RNA_t) + 1) #min normalise as well

SKCM_RNA_t_pair <- CPOP::pairwise_col_diff(SKCM_RNA_t)
MIA_RNA_t_pair <- CPOP::pairwise_col_diff(MIA_RNA_t)
identical(colnames(SKCM_RNA_t_pair), colnames(MIA_RNA_t_pair))


clean_names <- function(names){gsub("-", ".", names, fixed = TRUE)}
colnames(SKCM_RNA_t_pair) <- clean_names(colnames(SKCM_RNA_t_pair))
colnames(MIA_RNA_t_pair) <- clean_names(colnames(MIA_RNA_t_pair))

par(mfrow = c(1,2))
boxplot(t(SKCM_RNA_t_pair))
boxplot(t(MIA_RNA_t_pair))

SKCM = c(SKCM, 
         RNA_pair = t(SKCM_RNA_t_pair), 
         mapFrom = "RNASeq")

melanomaAssaysNorm = c(melanomaAssaysNorm, 
                       RNA_pair = t(MIA_RNA_t_pair), 
                       mapFrom = "NanoString")

SKCM@ExperimentList
melanomaAssaysNorm@ExperimentList
```


```{r}
par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["RNA_pair"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["RNA_pair"]]))
```

```{r}
summary(rowMeans(assay(SKCM@ExperimentList[["RNA_pair"]])))
summary(apply(assay(SKCM@ExperimentList[["RNA_pair"]]), 1, sd))
```


remove low sd RNA pairs
```{r}
RNA_pair <- assay(SKCM@ExperimentList[["RNA_pair"]])
keep_RNA_pair <- which(apply(RNA_pair, 1 , sd) > 0.1) ## soft filter
# keep_RNA_pair <- which(apply(RNA_pair, 1 , sd) > 0.5)
# keep_RNA_pair <- which(apply(RNA_pair, 1 , sd) > 1) ## hard filter
SKCM <- SKCM[list(RNA_pair = keep_RNA_pair),,]
melanomaAssaysNorm <- melanomaAssaysNorm[list(RNA_pair = keep_RNA_pair),,]
```

```{r}
par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["RNA_pair"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["RNA_pair"]]))
```


```{r}
assay(SKCM@ExperimentList[["RNA_pair"]]) <- scale(assay(SKCM@ExperimentList[["RNA_pair"]]))
assay(melanomaAssaysNorm@ExperimentList[["RNA_pair"]]) <- scale(assay(melanomaAssaysNorm@ExperimentList[["RNA_pair"]]))

par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["RNA_pair"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["RNA_pair"]]))
```







```{r}
#assay(SKCM@ExperimentList[["miRNA"]]) <- scale(assay(SKCM@ExperimentList[["miRNA"]]))
# assay(SKCM@ExperimentList[["RNASeq"]]) <- scale(assay(SKCM@ExperimentList[["RNASeq"]]))
#assay(melanomaAssaysNorm@ExperimentList[["microRNA"]]) <- scale(assay(melanomaAssaysNorm@ExperimentList[["microRNA"]]))
# assay(melanomaAssaysNorm@ExperimentList[["NanoString"]]) <- scale(assay(melanomaAssaysNorm@ExperimentList[["NanoString"]]))
```

```{r}
# par(mfrow = c(1,2))
# boxplot(assay(SKCM@ExperimentList[["RNASeq"]]))
# boxplot(assay(melanomaAssaysNorm@ExperimentList[["NanoString"]]))
```




# convert miRs to log ratios

```{r}
par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["miRNA"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["microRNA"]]))
```

```{r}
SKCM_miRNA_t <- t(assay(SKCM@ExperimentList[["miRNA"]]))
MIA_miRNA_t <- t(assay(melanomaAssaysNorm@ExperimentList[["microRNA"]]))
identical(colnames(MIA_miRNA_t), colnames(SKCM_miRNA_t))


SKCM_miRNA_t <- log2(SKCM_miRNA_t + 1)
MIA_miRNA_t <- log2(MIA_miRNA_t - min(MIA_miRNA_t) + 0.1) #min normalise as well

SKCM_miRNA_t_pair <- CPOP::pairwise_col_diff(SKCM_miRNA_t)
MIA_miRNA_t_pair <- CPOP::pairwise_col_diff(MIA_miRNA_t)
identical(colnames(SKCM_miRNA_t_pair), colnames(MIA_miRNA_t_pair))


clean_names <- function(names){gsub("-", ".", names, fixed = TRUE)}
colnames(SKCM_miRNA_t_pair) <- clean_names(colnames(SKCM_miRNA_t_pair))
colnames(MIA_miRNA_t_pair) <- clean_names(colnames(MIA_miRNA_t_pair))

par(mfrow = c(1,2))
boxplot(t(SKCM_miRNA_t_pair))
boxplot(t(MIA_miRNA_t_pair))

SKCM = c(SKCM, 
         miRNA_pair = t(SKCM_miRNA_t_pair), 
         mapFrom = "miRNA")

melanomaAssaysNorm = c(melanomaAssaysNorm, 
                       miRNA_pair = t(MIA_miRNA_t_pair), 
                       mapFrom = "microRNA")

SKCM@ExperimentList
melanomaAssaysNorm@ExperimentList
```


```{r}
par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["miRNA_pair"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["miRNA_pair"]]))
```

```{r}
summary(rowMeans(assay(SKCM@ExperimentList[["miRNA_pair"]])))
summary(apply(assay(SKCM@ExperimentList[["miRNA_pair"]]), 1, sd))
```


remove low sd miRs
```{r}
miRNA_pair <- assay(SKCM@ExperimentList[["miRNA_pair"]])
keep_miRNA_pair <- which(apply(miRNA_pair, 1 , sd) > 0.1) ## soft filter
# keep_miRNA_pair <- which(apply(miRNA_pair, 1 , sd) > 0.5)
# keep_miRNA_pair <- which(apply(miRNA_pair, 1 , sd) > 1) ## hard filter
SKCM <- SKCM[list(miRNA_pair = keep_miRNA_pair),,]
melanomaAssaysNorm <- melanomaAssaysNorm[list(miRNA_pair = keep_miRNA_pair),,]
```

```{r}
par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["miRNA_pair"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["miRNA_pair"]]))
```


```{r}
assay(SKCM@ExperimentList[["miRNA_pair"]]) <- scale(assay(SKCM@ExperimentList[["miRNA_pair"]]))
assay(melanomaAssaysNorm@ExperimentList[["miRNA_pair"]]) <- scale(assay(melanomaAssaysNorm@ExperimentList[["miRNA_pair"]]))

par(mfrow = c(1,2))
boxplot(assay(SKCM@ExperimentList[["miRNA_pair"]]))
boxplot(assay(melanomaAssaysNorm@ExperimentList[["miRNA_pair"]]))
```


rename mirna pairs

```{r}
head(rownames(assay(SKCM@ExperimentList[["miRNA_pair"]])))

```




# Train on SKCM 

```{r}
source("./MultiP-algm.R")
source("./MultiP-summary.R")

MultiAssayExperiment = SKCM

properties = c("TSA.retained", "Cost.Total")
properties_direction = c("high", "low") # best outcome of property
weights = c(0.5, 0.5)
confidenceCutoff = 0.1

# platformList = c("sampleInfo", "microRNA", "NanoString") 
# platformUnitCosts = c(50, 100, 100)

platformUnitCosts = c(100, 400, 200)
platformList = c("clinical", "RNA_pair", "miRNA_pair")

fixedPlatforms = 1
minPlatformSize = 5
performanceType = "Sample Error"

characteristics = NULL
seed=1
verbose=0
folds = 2
permutations = 50

crossValParams = CrossValParams(
    samplesSplits = "Permute k-Fold",
    permutations = 50,
    folds = 2,
    tuneMode = "none",
    parallelParams = SnowParam(workers = 2, RNGseed = seed)
)

modellingParams = ModellingParams(
    balancing = "none",
    transformParams  = NULL,
    selectParams = SelectParams("t-test"),
    trainParams = TrainParams("DLDA"),
    predictParams = PredictParams("DLDA"),
    doImportance = FALSE
)
```




# Main Code

```{r main, message = FALSE, warning = FALSE}
#filter NA's for class
classColumn = grep(classes, colnames(colData(MultiAssayExperiment)))
retainList = as.vector(!is.na(colData(MultiAssayExperiment)[classColumn]))
MultiAssayExperiment = MultiAssayExperiment[,retainList & !is.na(retainList),]

#Subset specified experiments from platformList 
MultiAssayExperiment = MultiAssayExperiment[,,platformList[platformList != "clinical"]]

#Subset complete cases only for all assays
MultiAssayExperiment = MultiAssayExperiment[,complete.cases(MultiAssayExperiment),]

dim(colData(MultiAssayExperiment))
```

```{r}
tic()
#set.seed(seed)
MultiPresults = MultiP.algm(MultiAssayExperiment,
                        platformList,fixedPlatforms,confidenceCutoff,
                        platformUnitCosts = platformUnitCosts , performanceType = performanceType,
                       classes = classes, crossValParams = crossValParams, modellingParams = modellingParams, characteristics = characteristics, minPlatformSize = minPlatformSize, seed = seed, verbose = verbose)
toc()
```

# Diagnostics

## Table
```{r}
summaryTable(MultiPresults)
```


## Bubble plot
```{r}
bubblePlot(MultiPresults, nVis = 8)
```


## Strata plot
```{r}
top_sequence = topBAccSequence(MultiPresults)
strataPlot(MultiPresults, top_sequence)
```


## Flow Chart
```{r}
top_sequence = topBAccSequence(MultiPresults)
flowChart(MultiPresults, top_sequence)
```


## Cohort Summary

```{r}
## convert numeric clinical variables back to categorical
# categorical_features = c"pathology_N_stage", "pathology_T_stage", "sex")
categorical_features = c("pathology_T_stage", "sex")
for(feature in categorical_features){
  colData(MultiPresults$MultiAssayExperiment)[,feature] = factor(colData(MultiPresults$MultiAssayExperiment)[,feature]-1)
}
```

### stage 1
```{r}
top_sequence = topBAccSequence(MultiPresults)
platform_of_interest = strsplit(top_sequence, "-")[[1]][1]
variables_of_interest = c("pathology_T_stage", "sex", "age")
cohortSummary(MultiPresults, top_sequence, platform_of_interest, variables_of_interest)
```


### stage 2
```{r}
platform_of_interest = strsplit(top_sequence, "-")[[1]][2]
cohortSummary(MultiPresults, top_sequence, platform_of_interest, variables_of_interest)
```


## Baseline

```{r}
for(platform in platformList){
results <-  modelsPredict(MultiAssayExperiment,
                         MultiPresults$generatedModelsResults,
                         platform,
                         classes, 
                         minPlatformSize,
                         verbose)
predict_class <- results$SSER$predict_class
true_class <- as.character(colData(MultiAssayExperiment)[,classes])
print(paste0("Platform: ", platform, 
             ", Accuracy:", round(mean(true_class == predict_class),3),
             ", Balanced Accuracy:", round(balanced_accuracy(true_class, predict_class),3)))
}

```

```{r}
top_sequence = topBAccSequence(MultiPresults)
nperm_SKCM = which(MultiPresults$TSER.summary$Platform.Sequence == top_sequence)[1]
stra_SKCM = MultiPresults$Results.stratification[[nperm_SKCM]] %>% filter(strata == "Retained")
classification_metrics(stra_SKCM$predict_class, stra_SKCM$true_class)

caret::confusionMatrix(data = factor(stra_SKCM$predict_class), reference = factor(stra_SKCM$true_class), positive = "Poor")
```



## apply model onto new data

```{r}
resultList = list()
```

```{r}
table(melanomaAssaysNorm$pathology_T_stage)
table(SKCM$pathology_T_stage)
```


```{r}
newData = DataFrame(colData(melanomaAssaysNorm)[,-1])
n_models = folds * permutations
newResults <- data.frame(sample = character(), class = character())


for(i in 1:n_models){
  result = modellingParams@predictParams@predictor(MultiPresults$generatedModels$clinical[[i]], newData) %>% suppressMessages
  newResults_temp <- data.frame(sample = rownames(result), class = result$class)
  newResults <- rbind(newResults, newResults_temp)
}

sampleIDs = rownames(colData(melanomaAssaysNorm))
pred_table <- table(newResults$sample, newResults$class)

class1 <- colnames(pred_table)[1]
class2 <- colnames(pred_table)[2]
class1_preds = pred_table[,1]
class1_preds = class1_preds[sampleIDs]
class2_preds = pred_table[,2]
class2_preds = class2_preds[sampleIDs]

preds <- data.frame(sampleID = sampleIDs, class1_pred = class1_preds, class2_pred = class2_preds)
preds <- preds %>% 
  mutate(predict_class = ifelse(class1_pred > class2_pred, class1, class2),
         true_class = as.character(colData(melanomaAssaysNorm)$class1vs4years)) %>% 
  mutate(sser = ifelse(class1 == true_class, class2_pred/n_models, class1_pred/n_models))
summary(preds$sser)
print(paste0("Accuracy:", round(mean(preds$true_class == preds$predict_class),3),
             ", Balanced Accuracy:", round(balanced_accuracy(preds$true_class, preds$predict_class),3)))

resultList[["clinical"]] = preds
```

```{r}
newData = DataFrame(t(melanomaAssaysNorm@ExperimentList[["RNA_pair"]]))
colnames(newData) <- paste0("RNA_pair_",colnames(newData))

n_models = folds * permutations
newResults <- data.frame(sample = character(), class = character())


for(i in 1:n_models){
  result = modellingParams@predictParams@predictor(MultiPresults$generatedModels$RNA_pair[[i]], newData) %>% suppressMessages
  newResults_temp <- data.frame(sample = rownames(result), class = result$class)
  newResults <- rbind(newResults, newResults_temp)
}

sampleIDs = rownames(colData(melanomaAssaysNorm))
pred_table <- table(newResults$sample, newResults$class)

class1 <- colnames(pred_table)[1]
class2 <- colnames(pred_table)[2]
class1_preds = pred_table[,1]
class1_preds = class1_preds[sampleIDs]
class2_preds = pred_table[,2]
class2_preds = class2_preds[sampleIDs]

preds <- data.frame(sampleID = sampleIDs, class1_pred = class1_preds, class2_pred = class2_preds)
preds <- preds %>% 
  mutate(predict_class = ifelse(class1_pred > class2_pred, class1, class2),
         true_class = as.character(colData(melanomaAssaysNorm)$class1vs4years)) %>% 
  mutate(sser = ifelse(class1 == true_class, class2_pred/n_models, class1_pred/n_models))
summary(preds$sser)
print(paste0("Accuracy:", round(mean(preds$true_class == preds$predict_class, na.rm = TRUE),3),
             ", Balanced Accuracy:", round(balanced_accuracy(preds$true_class, preds$predict_class),3)))


resultList[["RNA_pair"]] = preds
```


```{r}
newData = melanomaAssaysNorm@ExperimentList[["miRNA_pair"]] %>% assay %>% t %>% DataFrame
colnames(newData) <- paste0("miRNA_pair_",colnames(newData))

n_models = folds * permutations
newResults <- data.frame(sample = character(), class = character())


for(i in 1:n_models){
  result = modellingParams@predictParams@predictor(MultiPresults$generatedModels$miRNA_pair[[i]], newData) %>% suppressMessages
  newResults_temp <- data.frame(sample = rownames(result), class = result$class)
  newResults <- rbind(newResults, newResults_temp)
}

sampleIDs = rownames(colData(melanomaAssaysNorm))
pred_table <- table(newResults$sample, newResults$class)

class1 <- colnames(pred_table)[1]
class2 <- colnames(pred_table)[2]
class1_preds = pred_table[,1]
class1_preds = class1_preds[sampleIDs]
class2_preds = pred_table[,2]
class2_preds = class2_preds[sampleIDs]

preds <- data.frame(sampleID = sampleIDs, class1_pred = class1_preds, class2_pred = class2_preds)
preds <- preds %>% 
  mutate(predict_class = ifelse(class1_pred > class2_pred, class1, class2),
         true_class = as.character(colData(melanomaAssaysNorm)$class1vs4years)) %>% 
  mutate(sser = ifelse(class1 == true_class, class2_pred/n_models, class1_pred/n_models))
summary(preds$sser)
print(paste0("Accuracy:", round(mean(preds$true_class == preds$predict_class, na.rm = TRUE),3),
             ", Balanced Accuracy:", round(balanced_accuracy(preds$true_class, preds$predict_class),3)))

resultList[["miRNA_pair"]] = preds
```

## passing through tree

```{r}
set.seed(seed)
MultiPresults_predict <- MultiP.algm(melanomaAssaysNorm,
                      platformList = c("clinical", "miRNA_pair", "RNA_pair"), 
                      fixedPlatforms = 1,
                      confidenceCutoff,
                      platformUnitCosts = c("clinical" = 100, "miRNA_pair" = 400, "RNA_pair" = 200), 
                      performanceType = "Sample Error",
                      classes = "class1vs4years", 
                      crossValParams = crossValParams, 
                      modellingParams = modellingParams, 
                      characteristics = characteristics, 
                      minPlatformSize = 3, 
                      seed=seed, verbose=verbose, predict = TRUE, resultList = resultList)
```

# Diagnostics

## Table
```{r}
summaryTable(MultiPresults_predict)
```


## Bubble plot
```{r}
bubblePlot(MultiPresults_predict, nVis = 8)
```


## Strata plot
```{r}
top_sequence = topBAccSequence(MultiPresults)
strataPlot(MultiPresults_predict, top_sequence)
```


## Flow Chart
```{r}
top_sequence = topBAccSequence(MultiPresults)
flowChart(MultiPresults_predict, top_sequence)
```

## feature importance
```{r}
feat = platformList[1]
featureImportance(MultiPresults, feat, 20)

feat = platformList[2]
featureImportance(MultiPresults, feat, 20)

feat = platformList[3]
featureImportance(MultiPresults, feat, 20)
```

```{r}
library(ggraph)
library(igraph)

selected_features  <- MultiPresults$generatedModelsFeatures[["miRNA_pair"]] 

selected_features$node1 <- sapply(selected_features$Feature, function(x) strsplit(x, "--")[[1]][1])
selected_features$node2 <- sapply(selected_features$Feature, function(x) strsplit(x, "--")[[1]][2])

miRNA_pair_SKCM <- assay(MultiAssayExperiment@ExperimentList[["miRNA_pair"]])
miRNA_pair_SKCM <- miRNA_pair_SKCM[selected_features$Feature,] %>% t
CAD_inds <- colData(MultiAssayExperiment)$Outcome
poor_means <- colMeans(miRNA_pair_SKCM[CAD_inds == "Poor",])
good_means <- colMeans(miRNA_pair_SKCM[CAD_inds == "Good",])

# %>% 
#   mutate(CAD = colData(MultiAssayExperiment)$Outcome) %>%
#   group_by(CAD) %>%
#   summarise_all(mean)
# CAD_levels = miRNA_pair_SKCM_means$CAD
# miRNA_pair_SKCM_means <- miRNA_pair_SKCM_means %>% select(-CAD) %>% t %>% data.frame
# colnames(miRNA_pair_SKCM_means) <- CAD_levels

miRNA_pair_SKCM_means = data.frame(Good = good_means, Poor = poor_means)
rownames(miRNA_pair_SKCM_means) = colnames(miRNA_pair_SKCM)
miRNA_pair_SKCM_means <- mutate(miRNA_pair_SKCM_means, Diff = Good - Poor) %>% arrange(-abs(Diff)) %>% head(30)

miRNA_pair_SKCM_means$node1 <- sapply(rownames(miRNA_pair_SKCM_means), function(x) strsplit(x, "..", fixed = TRUE)[[1]][1])
miRNA_pair_SKCM_means$node2 <- sapply(rownames(miRNA_pair_SKCM_means), function(x) strsplit(x, "..", fixed = TRUE)[[1]][2])

# identical(rownames(miRNA_pair_SKCM_means), gsub("-", ".", selected_features$Feature, fixed = TRUE))
# selected_features$Diff <- miRNA_pair_SKCM_means$Diff
miRNA_pair_SKCM_means <- select(miRNA_pair_SKCM_means, c("node1", "node2", "Diff"))

graph <- graph_from_data_frame(miRNA_pair_SKCM_means)
ggraph(graph, layout = "linear", circular = TRUE) +
      ggraph::geom_edge_arc(aes(
        start_cap = label_rect(.data$node1.name),
        end_cap = label_rect(.data$node2.name),
        # width = .data$Freq,
        width = abs(.data$Diff),
        colour = factor(sign(.data$Diff), levels = c(1, -1), labels = c("Downregulated", "Upregulated"))))  +
      ggraph::scale_edge_colour_brewer(palette = "Set1", direction = -1, name="Direction") +
      ggraph::scale_edge_width_continuous(name = "Loading") +
      ggraph::geom_node_text(aes(label = sub("hsa.mir.", "", .data$name, fixed = TRUE)), size = 4.5) + 
      theme_graph(border = FALSE) #, plot_margin = ggplot2::margin(30, 50, 30, 50))

```


```{r}
selected_features  <- MultiPresults$generatedModelsFeatures[["RNA_pair"]] 

selected_features$node1 <- sapply(selected_features$Feature, function(x) strsplit(x, "--")[[1]][1])
selected_features$node2 <- sapply(selected_features$Feature, function(x) strsplit(x, "--")[[1]][2])

miRNA_pair_SKCM <- assay(MultiAssayExperiment@ExperimentList[["RNA_pair"]])
miRNA_pair_SKCM <- miRNA_pair_SKCM[selected_features$Feature,] %>% t
CAD_inds <- colData(MultiAssayExperiment)$Outcome
poor_means <- colMeans(miRNA_pair_SKCM[CAD_inds == "Poor",])
good_means <- colMeans(miRNA_pair_SKCM[CAD_inds == "Good",])


miRNA_pair_SKCM_means = data.frame(Good = good_means, Poor = poor_means)
rownames(miRNA_pair_SKCM_means) = colnames(miRNA_pair_SKCM)
miRNA_pair_SKCM_means <- mutate(miRNA_pair_SKCM_means, Diff = Good - Poor) %>% arrange(-abs(Diff)) %>% head(30)

miRNA_pair_SKCM_means$node1 <- sapply(rownames(miRNA_pair_SKCM_means), function(x) strsplit(x, "..", fixed = TRUE)[[1]][1])
miRNA_pair_SKCM_means$node2 <- sapply(rownames(miRNA_pair_SKCM_means), function(x) strsplit(x, "..", fixed = TRUE)[[1]][2])


miRNA_pair_SKCM_means <- select(miRNA_pair_SKCM_means, c("node1", "node2", "Diff"))

graph <- graph_from_data_frame(miRNA_pair_SKCM_means)
ggraph(graph, layout = "linear", circular = TRUE) +
      ggraph::geom_edge_arc(aes(
        start_cap = label_rect(.data$node1.name),
        end_cap = label_rect(.data$node2.name),
        # width = .data$Freq,
        width = abs(.data$Diff),
        colour = factor(sign(.data$Diff), levels = c(1, -1), labels = c("Downregulated", "Upregulated"))))  +
      ggraph::scale_edge_colour_brewer(palette = "Set1", direction = -1, name="Direction") +
      ggraph::scale_edge_width_continuous(name = "Loading") +
      ggraph::geom_node_text(aes(label = .data$name), size = 3) + 
      theme_graph(border = FALSE) #, plot_margin = ggplot2::margin(30, 50, 30, 50))

```

## Cohort Summary

```{r}
## convert numeric clinical variables back to categorical
categorical_features = c("pathology_T_stage", "sex")
for(feature in categorical_features){
  colData(MultiPresults_predict$MultiAssayExperiment)[,feature] = factor(colData(MultiPresults_predict$MultiAssayExperiment)[,feature]-1)
}
```

### stage 1
```{r}
top_sequence = topBAccSequence(MultiPresults)
platform_of_interest = strsplit(top_sequence, "-")[[1]][1]
cohortSummary(MultiPresults_predict, top_sequence, platform_of_interest, variables_of_interest)
```


### stage 2
```{r}
platform_of_interest = strsplit(top_sequence, "-")[[1]][2]
cohortSummary(MultiPresults_predict, top_sequence, platform_of_interest, variables_of_interest)
```

```{r}
top_sequence = topBAccSequence(MultiPresults)
nperm_MIA = which(MultiPresults_predict$TSER.summary$Platform.Sequence == top_sequence)[1]
stra_MIA = MultiPresults_predict$Results.stratification[[nperm_MIA]] %>% filter(strata == "Retained")
classification_metrics(stra_MIA$predict_class, stra_MIA$true_class)
caret::confusionMatrix(data = factor(stra_MIA$predict_class), reference = factor(stra_MIA$true_class), positive = "Poor")
```


## DLDA Importance plot

```{r}
DLDAloadings(MultiPresults, platformList[1])
DLDAloadings(MultiPresults, platformList[2])
DLDAloadings(MultiPresults, platformList[3])
```



```{r}
sessionInfo()
```

